A Compound Decision Approach to Covariance Matrix Estimation

### 2021-10-02 response_reviews.tex

\documentclass[12pt]{article}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{comment}
\usepackage{lscape}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\def\bs{\boldsymbol}
\def\bb{\mathbb}

\title{}

\begin{document}

\section{Comments to the editor}
\begin{enumerate}
\item \emph{As a journal, Biometrics presents papers that focus on the development of new methods and results of use in biosciences. As of now, the application seems more an afterthought, the use of a convenient (easily available) dataset to fit the scope of the journal. I believe that the authors should motivate their approach from the gene expression data or other bio-related applications starting from the Introduction.}
\item \emph{It is not really intuitively clear why the vectorization works, despite the approach not taking into consideration the structure of the covariance matrix. Also, the manuscript proposes a series of ad-hoc expedients to implement the method. See the clustering-based exemplar algorithm and replacing non-positive eigenvalues with a chosen positive value $c$ in the positive-definite correction. These expedients do not seem to be completely justified by the proposed separable compound-decision theory framework, and their effects on the estimation are unclear. A referee also commented on the impossibility to follow closely the implementation of the EM algorithm. In short, while I appreciate that the theoretical justification for the asymptotic efficiency of the proposed vectorization is not clear, the current implementation reads more like a clever algorithm than a well-founded statistical method. The authors should justify all the steps of the algorithm and provide an assessment of their contribution to its overall performance.}

\item \emph{I agree with one of the referees that the title should be re-thought, since compound decision theory or compound decisions appear quite relevant to the approach, more than the empirical Bayes aspect of the paper.}\\
We agree that compound decision theory should be more emphasized than empirical Bayes in our paper. Following your suggestions, we have changed the title from "A Empirical Bayes Approach to Covariance Matrix Estimation" to "A Compound Decision Approach to Covariance Matrix Estimation".

\item \emph{on page 4, it appears to me that the choice of the risk function $\hat{R}$ does not correspond to the Frobenius risk. The rationale for choosing this risk function should be provided. It somehow appears the choice is dictated by Ledoit et al (2004) and Proposition 1, and not by independent first principles. If so, it should be explicitly stated before.}\\
We squared and scaled the Frobenius norm $\|A\|_F = \sqrt{\sum_{j,k=1}^pa_{jk}^2}$ to $\frac{1}{p^2}\sum_{j,k=1}^p a_{jk}^2$ in order to apply fundamental theorem of compound decisions, for which the risk function should be scaled by the number of parameters because the empirical distribution has mass probability of $\frac{1}{d}$ where $d$ is the number of parameters.\\
The transformation of Frobenius norm will not change the value of $(\widehat{\beta}_S,\widehat{\beta}_I)$ since it is determined by minimizing $RE(\beta_S,\beta_I)$.
\item \emph{Figures 1 and 2 do not allow to appreciate uncertainty. Perhaps a table would be better.}\\
We add the tables of the medians and uncertainty of Frobenius errors in supporting information. The uncertainty of Frobenius error is measured by the difference of 75\% quantile and 25\% quantile in 200 replicates. We use this range to measure the uncertainty because the errors of most methods are not following normal distribution. 
\item \emph{The manuscript requires careful proofreading and improvements. in the structure of the presentation and the visualization of the results. Non-exhaustive examples are reported below:}

\emph{- p.1 “To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrix” $\to$ matrices}

It has been fixed.

\emph{- p. 1 “about the structure of population covariance matrix” $\to$ of the population covariance matrix}

It has been fixed.

\emph{- p. 2 “Other common structured methods assume that the covariance matrix is banding” $\to$ banded?}

It has been edited to 'banded'.

\emph{- p. 2 “yet can still outperform the sample covariance matrix” What does this mean?}

It means that these unstructured model have been shown to have lower estimation risk comparing to sample covariance matrix. For example, in \cite{ledoit2004}, the numerical result shows that their linear shrinkage method has much lower risk than sample covariance matrix. To make it clear, we have changed to "yet still have less estimation error comparing to the sample covariance matrix".

\emph{- p. 3 “is squared error loss”$\to$ is the squared error loss}

It has been fixed.

\emph{- p. 4 “Now consider the problem of estimating the vectorized $\boldsymbol{\Sigma}$ under risk 1” $\to$ risk (1) Equations are typically referenced between brackets. The same issue appears in many other points where equations are referenced throughout the manuscript.}

We used to use \textbackslash ref to reference equation, we have changed it to \textbackslash eqref so that they are referenced between brackets.

\emph{- p. 8 “A major advantage of nonparametric estiation of the prior”$\to$ estimation}

It has been fixed.

\emph{- page 8, “Figure 1 shows that different $K$ have similar estimation accuracy compared to the exemplar algorithm, while Table 1 shows that they can be significantly faster” However, at this point Figure 1 refers to models that have not yet been presented. Also, the reference to Table 1 anticipates point made in Section 3.2}

We have moved the setting of the experiment to the start of Section 3 Numerical Results.

\emph{- p. 16 “Finally, we have so far assumed that our data multivariate normal” $\to$ are normal}

It has been fixed.

\emph{- In Equations (1) and (2) the arguments of the risk functions are $\theta$ and $\delta$. However, on page 5 the arguments of $\hat{R}$ are $\beta_{S}$, and $\beta_{I}$.}

We have changed the notation $\hat{R}$ to $\text{RE}$ which means risk estimator and this differentiates it from $R$. 
\end{enumerate}

\section{Comments to Referee1}
Thank you so much for the detailed and inspiring suggestions on our paper. We learnt a lot through them. We agree with your comments and address them in the updated draft version. Please find the following responses for more details.
\subsection{comments to section "CorShrink"}
After carefully reviewed \cite{dey2018}, we think it is very related to our paper and it is valuable to make the comparison.

As mentioned in Section 1, our method is quite similar to \cite{dey2018}. Both of us used empirical Bayes estimator and nonparametric estimation of prior distribution. The difference is that our method aim to estimate covariance matrix which requires to also shrink standard deviations.

We also compared our two methods by numerical experiments which are shown in supporting information. We designed two simulations to compare the performance on correlation matrix estimation and covariance matrix estimation. In the first simulation, we compare covariance matrix estimation. For corShrink, $\widehat{\bs{\Sigma}} = \text{diag}(\widehat{SD}) \widehat{\bs{R}} \text{diag}(\widehat{SD})$, where $\widehat{SD}$ are sample standard deviations. In the second simulation, $\bs{\widehat{R}} = \text{diag}(1/\widehat{SD})\widehat{\bs{\Sigma}}\text{diag}(1/\widehat{SD})$, where $\widehat{SD}$ are standard deviations derived from $\widehat{\bs{\Sigma}}$. For each setting, we take $n$ as 20 and 100.

\subsection{comments to section "Separable rules, symmetry and the diagonal"}
\begin{equation}\label{R1}
\widehat{\Sigma} = (\widehat{\sigma}_{jk})_{1\leq j,k\leq p},\quad \widehat{\sigma}_{jk} = t(X_{\cdot,j}, X_{\cdot,k}), \quad t:\bb{R}^n\times \bb{R}^n\to \bb{R}
\end{equation}
\begin{enumerate}
\item \emph{While reading the paper I was a bit confused by the fact that (R1) treats diagonal entries and off-diagonal in the same way. It would feel more natural to me to define a separable estimator as in (R1) for $j\neq k$, and as follows on the diagonal:}
$$\hat{\sigma}_{jj} = \tilde{t}(X_{\cdot j}),\quad \tilde{t}:\bb{R}^n \to \bb{R}$$
We have changed the rule to 
\begin{equation}
  \label{separable}
  \mathcal{S} = \{\bs{\delta} : \delta_{kj} = \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p,\quad 
   \delta_{jj} = \widetilde{t}(\bs{X}_{\cdot j}),\,j=1,\ldots,p)\},
\end{equation}
\emph{In the Multivariate Gaussian case this is not an issue, since $X_{\cdot j}\neq X_{\cdot k}$ for $j\neq k$ with probability 1. But the proposed approach would also apply with minor modifications to e.g., discrete measurements (with known likelihood) in which case $X_{\cdot j}\neq X_{\cdot k}$ could happen with non-zero probability.} \\
We agree that it is not very precise to use $I(\bs{X}_{\cdot j}=\bs{X}_{\cdot k})$ to differentiate on-diagonal entries and off-diagonal entries. So we changed separable rule to the above class of rules.

\emph{Furthermore, while this issue is ignored for most of the paper, diagonal and off-diagonal entries are treated in a different way at the end of Section 2.3 (in the context of the clustering based algorithm). I believe it would be valuable to make the above distinction more explicit!}
\item \emph{A related question I had concerns the symmetry of the estimator. Does the estimator satisfy the symmetry property $\hat{\sigma}_{jk} =\hat{\sigma}_{kj}$? }

Yes, we only calculate estimates of lower triangular matrix of $\Sigma$, i.e., $\hat{\sigma}_{jk}$ where $j\geq k$. Then the upper triangular matrix estimation is calculated by the symmetry of covariance matrix $\sigma_{kj} = \sigma_{jk}$. Correspondingly, we should fix the Bayesian model in Section2.2 from
\begin{equation}
  \label{bayesian}
  \bs{A} \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
  \quad
  \bs{\eta} \sim G_p(a,b, \gamma) =  \frac{1}{p^2} \sum_{j,k=1}^p I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
\end{equation}
to 
\begin{equation}
  \label{bayesian nondigaonal}
 (\bs{X}_{\cdot j},\bs{X}_{\cdot k})\mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
  \quad
  \bs{\eta} \sim G_{nd}(a,b, \gamma) =  \frac{2}{p(p-1)} \sum_{1\leq k< j \leq p} I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
\end{equation}
and
\begin{equation}
  \label{bayesian nondigaonal}
  \bs{X}_{\cdot j} \mid a \sim \widetilde{f}(\cdot \mid a),
  \quad
  a \sim G_{d}(a) =  \frac{1}{p} \sum_{j=1}^p I(\sigma_j \leq a).
\end{equation}
, as well as the proposed estimator from
\begin{equation}
  \label{Gp hat}
  \widehat{G}_p = \argmax_{G \in \mathcal{G}} \prod_{j,k = 1}^p \int f(\bs{A}_{jk} \mid \bs{\eta}) dG(\bs{\eta}),
\end{equation}
to
\begin{equation}
  \label{Gp hat revision}
  \widehat{G}_{nd} = \argmax_{G_{nd} \in \mathcal{G}} \prod_{1\leq k< j \leq p} \int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) dG_{nd}(\bs{\eta})\prod_{j=1}^p \int \widetilde{f}(\bs{X}_{\cdot j} \mid a) dG_{d}(a),
\end{equation}
\emph{Similarly, the reader may wonder whether the estimation procedure enforces symmetry in the prior $G$ in (6) with respect to $\sigma_j,\sigma_k$.}

In correspondence, we modified our sample exemplar algorithm, we first clusters sample points $(s_j,s_k, r_{jk})$ of lower triangular part of sample covariance matrix and get the $K$ centroids $(a,b,\gamma)$, then find their symmetric points $(b,a,\gamma)$. We also updated EM algorithm to guarantee each $(a,b,\gamma)$ and $(b,a,\gamma)$ has the same weight. 
\item \emph{The paper emphasizes the Multivariate Gaussian problem. In that case, the sample covariance is sufficient, and so, it would seem more natural to me to define a separable rule, as one that satisfies,
$$\hat{\sigma}_{jk} = \bar{t}(X_{\cdot j}^TX_{\cdot k}),\quad \bar{t}: \bb{R}\to \bb{R}_{+}$$
instead of (R1). I am wondering why the authors chose to use (R1) instead.}

$X_{\cdot j}^TX_{\cdot k}$ is sufficient statistics of the single entry $\sigma_{jk}$. In fact, for $(\bs{X}_{\cdot,j}, \bs{X}_{\cdot,k})$, the sufficient statistics for $\eta$ is $\bs{S}_{jk} = \begin{pmatrix} s_{jj}&s_{jk} \\  s_{kj}&s_{kk} \end{pmatrix}$. We also tried density based on single entries $s_{jk}$ before but the performance is not as good as using $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$.
\item \emph{If I understand correctly, one of the benefits of estimator (3) is that it works without specific assumptions on the noise distribution. Instead the approach here appears to more heavily rely on a known likelihood (and specifically a multivariate Gaussian likelihood). It would be great if some simulations could be conducted under misspecification, wherein the data is not multivariate Gaussian, but the approach is applied assuming multivariate Gaussian data.}

This is a good point. To measure the performance of our method on nonGaussian distribution, we did some simulations on negative binomial distribution. We first generate independent distributed $Y$ and then let $X = L Y$, where L is the Cholesky decomposition of $\Sigma$. The simulation results are presented in supporting information.
\end{enumerate}

\subsection{comments to section "Compound decision theory or empirical Bayes?"}
\begin{enumerate}
\item \emph{The result following equation (6) could be presented separately as a Theorem (or Proposition). It should be highlighted that part of the statement is that the estimator in (R1) that optimizes (R2) depends on all the $\sigma_{j}$ and $\rho_{jk}$.}

We rewrote the equation after (6) as a separate proposition to show the equivalence between the Frobenius risk and Bayes risk.
\item \emph{The term “compound decision theory” or “compound decisions” etc. deserves to be in the title!}

Yes, we agree that "compound decision" is more deserved to be highlighted in the paper so we replaced "nonparametric empirical Bayes" with "compound decision" in the title.
\item \emph{On the other hand, I would downemphasize the empirical Bayes aspect of the paper (and my preference would be to even remove it from the title). To me, an empirical Bayes analysis (especially since here the goal is to perform well, in a frequentist sense, in terms of (R2)) has the following flavor: “Let us mimic what a Bayesian would do, when we do not know the true prior, and in a way that we can get frequentist guarantees as well”. However, I do not think a Bayesian would start approaching this problem by first positing (6) and then using (8)2. Instead a more natural Bayesian approach would be to set a prior for the full covariance matrix $\Sigma$, e.g., as in Berger et al. [2020] and references therein. I would restrict the discussion of empirical Bayes to just saying that model (6) appears formally when optimizing over separable rules (R1). To approximate the unknown optimal rule, one follows an empirical Bayesian “Generalized Maximum Likelihood” approach with respect to the pairwise composite likelihood.3}
\end{enumerate}

\subsection{comments to section "other remarks"}
\begin{enumerate}
\item \emph{It would be of great service to the community if the authors could provide code implementing the method and reproducing the results!}

We put our code of methods and simulation in supporting information.
\item \emph{Related to the above point, based on the description in the manuscript it would not be possible for me to fully reproduce the results. For example, does the actual implementation use the EM algorithm or an interior point convex solver? If the EM algorithm is used, how many steps are taken or what is the termination criterion? And why is the EM algorithm used (since for these empirical Bayes problem it typically converges extremely slowly compared to interior point solvers, as carefully demonstrated by Koenker and Mizera [2014])? On the other hand, if an interior point solver is used, which one?}

In our implementation, we used EM algorithm to estimate the prior distribution. We added the related details to Section 3 Numerical Results. The maximum number of iterations is set as 200. Iteration early stops when the criterion $\frac{L(\bs{w}^{(k+1)}) - L(\bs{w}^{(k)})}{L(\bs{w}^{(k)})}\leq 10^{-4}$ is met. We also tried interior point, but it always failed to find the optimal solution. The possible reason is the density $f(\cdot|a,b,\gamma)$ has very different scales, from $10^{-22}$ to $2$ and this makes interior point method difficult to solve.
\item \emph{First paragraph of page 3, “they modify only the sample eigenvalues and not the sample eigenvectors”: The situation here is a bit more nuanced, since e.g., sometimes using the sample eigenvectors is provably the optimal thing to do (e.g., if one constrains themselves to rotationally invariant estimators as in Bun et al. [2016] and references therein)}

The sentence "However, they modify only the sample eigenvalues and not the sample eigenvectors." is not very appropriate so we changed it to "However, as rotation-equivariant estimators (Stein, 1975, 1986), shrinkage estimators modify only the sample eigenvalues and keep the sample eigenvectors.".
\item \emph{Page 5, “it is straightforward to show that”: The “straightforward” could be replaced by a short proof in the appendix.}

We added the proof in the appendix.

\emph{Also could it be that the risk estimate is not exactly unbiased (because of dividing by $n$ and not $n-1$ in the definition of $\hat{\Delta}_{jk}^2$)?}
You're right. We realized that the previous defined $\hat{\Delta}_{jk}^2$ is not unbiased estimator of Frobenius risk of $\bs{\delta}$ (1). So we changed the text to describe it as an "approximately unbiased risk estimator" since the factor $\frac{n-1}{n}$ is closed to 1. It is also shown in appendix B.

\emph{Where is the sample covariance $\bs{S}$ defined?}

In the updated manuscript, $\bs{S}$ is defined as $\bs{S} = \frac{1}{n}\bs{X}^\top \bs{X}$ as it is first mentioned in page 4, section 2.1.

\item \emph{Section 2.5, “Projections in terms of other matrix norms are also possible”. Is this actually being done in the implementation?}

No. We did not implement it in terms of other norms.

\emph{Else perhaps remove this sentence.}

### 2021-11-02 msg.tex
edited intro and compound decision formulation

% \documentclass[12pt]{article}
\documentclass[useAMS,referee,usenatbib]{biom}
%\usepackage[utf8]{inputenc}
%\usepackage{amsmath, authblk}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{comment}
\usepackage{lscape} 
%\usepackage{graphics}
%\usepackage{afterpage}
%\usepackage{xcolor}
%\usepackage{citeppp}
%\usepackage[usenatbib]{natbib}
%\usepackage[round]{natbib}
%\usepackage{url} 
\usepackage{graphicx}
%\usepackage{subcaption}

@@

\title[Empirical Bayes Covariance Matrix Estimation]{A Compound Decision Approach to Covariance Matrix Estimation}

@@

\label{firstpage}
\begin{abstract}
  We propose an empirical Bayes method to estimate high-dimensional covariance matrices. Our procedure centers on vectorizing the covariance matrix and treating matrix estimation as a vector estimation problem. Drawing from the compound decision theory literature, we introduce a new class of decision rules that generalizes several existing procedures. We then use a nonparametric empirical Bayes $g$-modeling approach to estimate the oracle optimal rule in that class. This allows us to let the data itself determine how best to shrink the estimator, rather than shrinking in a pre-determined direction such as toward a diagonal matrix. Simulation results and a gene expression network analysis shows that our approach can outperform a number of state-of-the-art proposals in a wide range of settings, sometimes substantially.
\end{abstract}

@@

Covariance matrix estimation is a fundamental statistical problem that plays an essential role in various applications, such as portfolio management \citep{ledoit2003improved}, genomics \citep{schafer2005shrinkage}, and array signal processing \citep{abramovich2001locally}. However, in modern problems the number of features can be of the same order as or exceed the sample size, and the standard sample covariance matrix estimator behaves poorly in this regime. To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrces. These can roughly be divided into two groups, according to whether they impose assumptions of the structure of population covariance matrix.

Structured methods make structural assumptions about the population covariance matrix. One class models the population covariance matrix as sparse. The most common method to address this problem is thresholding \citep{rothman2009generalized, cai2011adaptive}. Penalized likelihood methods \citep{xue2012positive} can also estimate large-scale sparse covariance matrix by penalizing a log-likelihood function. Another class of methods assume the data arise from a factor model \citep{fan2008high}, so that the covariance matrix has low intrinsic dimension. Other common structured methods assume that the covariance matrix is banded \citep{li2017estimation} or Toeplitz matrix \citep{liu2017covariance}.

In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet still have less estimation error comparing to the sample covariance matrix. A first example was the linear shrinkage approach of \citet{ledoit2004well}, which shrinks the sample covariance matrix toward a scaled identity matrix. More recently, nonlinear shrinkage methods were developed \citep{ledoit2012nonlinear, ledoit2019quadratic, lam2016nonparametric}. These shrink the eigenvalues of the sample covariance matrix toward clusters. Linear shrinkage can be viewed as a special case of nonlinear shrinkage, as it shrinks sample eigenvalues toward their global mean.

Nonlinear shrinkage estimators have desirable optimality properties \citep{ledoit2018analytical} and show excellent performance. However, as the class of rotation-equivariant estimators \citep{stein1975estimation, stein1986lectures}, nonlinear shrinkage estimators modify only the sample eigenvalues and keep the sample eigenvectors. It is known that sample eigenvectors are not consistent estimators of population eigenvectors when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}. This suggests that there may exist a class of unstructured estimators that can outperform nonlinear shrinkage.

%In this paper, we propose a new covariance matrix estimator which is approximately optimal among all separable estimators. Our method is motivated by the similarity between covariance estimation and vector mean estimation. Covariance matrix estimation under Frobenius loss can be viewed as a vector estimation problem by converting the matrix to a vector. \citep{cai_liu} used this motivation to generalize thresholding method in mean vector estimation to large sparse covariance matrix estimation. Nevertheless, they are not totally the same because covariance matrix has repeated entries due to its symmetry. In mean vector estimation problem, optimality has been achieved among all separable decision rules. \citep{jiang_zhang2009} and \citep{brown_greenshtein} proposed empirical Bayes methods to approximate the optimal separable estimator. We proposed to apply vector estimation methods to covariance estimation problem and expect that these methods will also have good properties here. Result in simulation and real data set shows that our estimator has good performance in different scenarios.

Here we propose a new unstructured estimator for high-dimensional covariance matrices. Our approach centers on vectorizing the covariance matrix and treating matrix estimation as a vector estimation problem. We do this because it allows us to use a nonparametric empirical Bayes shrinkage procedure, which has been shown in the compound decision literature to have excellent properties \citep{jiang2009general, koenker2014convex}. We then reassemble the estimated vector into matrix form and project onto the space of positive-definite matrices to give our final estimator. Surprisingly, though our vectorized approach essentially ignores the matrix structure, it can still substantially outperform a number of state-of-the-art proposals in simulations and a real data analysis. Our approach is similar to \citet{dey2018corshrink}, which also applied empirical Bayes shrinkage method to estimate correlation matrix. Different from our method, their corShrink method has different target and they do not shrink standard deviations.

The article is organized as follows. In Section \eqref{sec:method}, we briefly review compound decision theory and then introduce our proposed approach. In Section \eqref{numerical results} we illustrate the performance of our method in simulations and a gene expression dataset. Finally, Section \eqref{sec:discussion} concludes with a discussion. Our procedure is implemented in the R package \verb|cole|, available on GitHub.

\section{\label{sec:method}Method}
\subsection{\label{sec:compound}Compound decision problem formulation}
Suppose we have $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$. The purpose of this paper is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$ that minimizes the scaled squared Frobenius risk
\begin{equation}
\label{frobenius risk}
R(\bs{\Sigma}, \bs{\delta}) = \frac{1}{p^2} \sum_{j,k=1}^p \mathbb{E}[\{\delta_{jk}(\bs{X})-\sigma_{jk}\}^2],
\end{equation}
where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate.

Our proposed approach is motivated by two observations. First, \eqref{frobenius risk} shows that estimating $\bs{\Sigma}$ under Frobenius risk is equivalent to simultaneously estimating every component of the vector $(\sigma_{11},\ldots,\sigma_{pp})^\top$ under a loss function that aggregates errors across components. Second, this type of vector estimation problem has been well-studied in the compound decision literature. Thus, recent advances in vector estimation may be profitably applied to covariance matrices.

We first briefly review compound decision problems. Introduced by \citet{robbins1951asymptotically}, these problems study the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$. Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk
\begin{equation}
\label{compound risk}
R(\bs{\theta},\bs{\delta}) = \frac{1}{n}\sum_{i=1}^n \bb{E} L(\theta_i,\delta_i(\bs{Y}))
\end{equation}
where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$. A classical example is the homoscedastic Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is the squared error loss \citep{johnstone2017gaussian}.

A key property of compound decision problems is that while a given $Y_i$ seems to offer no information about any specific $\theta_j$ when $j \ne i$, borrowing information across all components of $\bs{Y}$ to estimate $\bs{\theta}$ is superior to estimating each $\theta_i$ using the corresponding $Y_i$ alone. A classical example of this phenomenon is the James-Stein estimator \citep{james1961estimation}, which estimates $\bs{\theta}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$. It is well-known that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$. A long line of subsequent work has led to much more sophisticated and accurate procedures for estimating $\bs{\theta}$ \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}.

>>>>

% \documentclass[12pt]{article}
\documentclass[useAMS,referee,usenatbib]{biom}
%\usepackage[utf8]{inputenc}
%\usepackage{amsmath, authblk}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{comment}
\usepackage{lscape} 
%\usepackage{graphics}
%\usepackage{afterpage}
%\usepackage{xcolor}
%\usepackage{citeppp}
%\usepackage[usenatbib]{natbib}
%\usepackage[round]{natbib}
%\usepackage{url} 
\usepackage{graphicx}
% \usepackage{subcaption}
\usepackage{xcolor}

@@

\def\blue{\textcolor{blue}}

\title[Coumpound Decision Covariance Matrix Estimation]{A Compound Decision Approach to Covariance Matrix Estimation}

@@

\label{firstpage}
\begin{abstract}
  Covariance matrix estimation is a fundamental statistical task in many applications, but the sample covariance matrix is sub-optimal when the sample size is comparable to or less than the number of features. Such high-dimensional settings are common in modern genomics, where covariance matrix estimation is frequently employed as a method for inferring gene networks. To achieve estimation accuracy in these settings, existing methods typically either assume that the population covariance matrix has some particular structure, for example sparsity, or apply shrinkage to better estimate the population eigenvalues. In this paper, we propose a new approach to estimating high-dimensional covariance matrices that can have better performance than existing methods. We first frame covariance matrix estimation as a compound decision problem. This motivates defining a class of decision rules and using a nonparametric empirical Bayes $g$-modeling approach to estimate the optimal rule in the class. Simulation results and gene network inference in a single cell RNA-seq dataset show that our approach is comparable to or can outperform a number of state-of-the-art proposals in a wide range of settings.
\end{abstract}

@@

Covariance matrix estimation is a fundamental statistical problem that plays an essential role in various applications. However, in modern problems the number of features can be of the same order as or exceed the sample size. This high-dimensional setting is especially common in genomics, where covariance matrices are used to model gene networks but the number of genes can be much larger than the number of biological replicates \citep{schafer2005shrinkage, markowetz2007inferring}. Accurate gene network estimation has become even more important with the advent of single-cell RNA-sequencing (scRNA-seq) technology, where interest now centers on inferring cell- or cell type-specific gene networks \citep{aibar2017scenic, dai2019cell, stegle2015computational}. For example, in Section \ref{gene analysis} we study cell type-specific gene network inference using scRNA-seq data from honey bee brains. \blue{Need to do analysis}

A common approach to estimating gene expression networks is to use the standard sample covariance or corrlation matrices \citep{langfelder2008wgcna, zhang2005general}. However, these estimators behave poorly in the high-dimensional regime. To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrices. These can roughly be divided into two groups, according to whether they impose assumptions of the structure of population covariance matrix.

Structured methods make assumptions about the form of the population covariance matrix. One popular class of methods assumes that the matrix is sparse, or has many zero entries. The most common strategy in this class is to threshold the entries of the sample covariance \citep{rothman2009generalized, cai2011adaptive}, but penalized likelihood methods \citep{xue2012positive} have also been used. A second class of methods assume the data arise from a factor model \citep{fan2008high}, so that the covariance matrix has low intrinsic dimension. Other common structured methods assume that the covariance matrix is banded \citep{li2017estimation} or Toeplitz \citep{liu2017covariance}.

In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet still have lower estimation error than the sample covariance matrix. A first example was the linear shrinkage approach of \citet{ledoit2004well}, which shrinks the sample covariance matrix toward a scaled identity matrix. More recently, nonlinear shrinkage methods were developed \citep{ledoit2012nonlinear, ledoit2019quadratic, lam2016nonparametric}. These shrink the eigenvalues of the sample covariance matrix in a data-adaptive fashion. Linear shrinkage can be viewed as a special case of nonlinear shrinkage, as it shrinks sample eigenvalues toward their global means.

Nonlinear shrinkage estimators have desirable optimality properties \citep{ledoit2018analytical} and show excellent performance. However, most existing nonlinear shrinkage estimators belong to the class of rotation-equivariant estimators \citep{bun2016rotational, stein1975estimation, stein1986lectures}, which modify the sample eigenvalue estimates but still use the sample eigenvectors. On the other hand, sample eigenvectors are not consistent when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}. This suggests that certain classes of non-rotationally invariant estimators may outperform existing unstructured covariance matrix estimators.

% In this paper, we propose a new covariance matrix estimator which is approximately optimal among all separable estimators. Our method is motivated by the similarity between covariance estimation and vector mean estimation. Covariance matrix estimation under Frobenius loss can be viewed as a vector estimation problem by converting the matrix to a vector. \citep{cai_liu} used this motivation to generalize thresholding method in mean vector estimation to large sparse covariance matrix estimation. Nevertheless, they are not totally the same because covariance matrix has repeated entries due to its symmetry. In mean vector estimation problem, optimality has been achieved among all separable decision rules. \citep{jiang_zhang2009} and \citep{brown_greenshtein} proposed empirical Bayes methods to approximate the optimal separable estimator. We proposed to apply vector estimation methods to covariance estimation problem and expect that these methods will also have good properties here. Result in simulation and real data set shows that our estimator has good performance in different scenarios.

Here we propose a new class of unstructured non-rotationally invariant estimators for high-dimensional covariance matrices. Our approach is based on interpreting the covariance matrix estimation problem as a compound decision problem \citep{robbins1951asymptotically}; see Section \ref{sec:compound}. This framing motivates us to vectorize the covariance matrix and solve the resulting vector estimation problem using a nonparametric empirical Bayes procedure, which has been shown in the compound decision literature to have excellent properties \citep{jiang2009general, koenker2014convex, efron2019bayes}. We then reassemble the estimated vector into matrix form and project onto the space of positive-definite matrices to give our final estimator. Surprisingly, though our vectorized approach essentially ignores the matrix structure, it can still substantially outperform a number of state-of-the-art proposals in simulations and a real data analysis. Our compound decision approach is conceptually similar the empirical Bayes approach taken by \citet{dey2018corshrink} to estimate a high-dimensional correlation matrix. We discuss in detail the connections with their work in Section \ref{sec:proposed}. \blue{Need to add to numerical results}

The article is organized as follows. In Section \eqref{sec:method}, we briefly review compound decision theory and then introduce our proposed approach. In Section \eqref{numerical results} we illustrate the performance of our method in simulations and a gene expression dataset. Finally, Section \eqref{sec:discussion} concludes with a discussion. Our procedure is implemented in the R package \verb|cole|, available on GitHub. \blue{Is this implemented in the package yet?}

\section{\label{sec:compound}Compound decision problem formulation}
\subsection{\label{sec:background}Background}

Given $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$, our goal is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$. A common measure of estimation performance is the scaled squared Frobenius risk
\begin{equation}
\label{frobenius risk}
R(\bs{\Sigma}, \bs{\delta}) = \frac{1}{p^2} \sum_{j,k=1}^p \mathbb{E}[\{\delta_{jk}(\bs{X})-\sigma_{jk}\}^2],
\end{equation}
where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate. This paper constructs a new non-rotationally invariant $\bs{\delta}$ that in simulations and a data analysis has comparable or lower risk \eqref{frobenius risk} compared to existing methods.

Our proposed approach is motivated by the observation that minimizing \eqref{frobenius risk} is a type of compound decision problem, which we briefly review in this section. Introduced by \citet{robbins1951asymptotically}, compound decision problems study the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$. Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk
\begin{equation}
\label{compound risk}
R(\bs{\theta},\bs{\delta}) = \frac{1}{n}\sum_{i=1}^n \bb{E} L(\theta_i,\delta_i(\bs{Y}))
\end{equation}
where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$. A classical example is the homoscedastic Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is the squared error loss \citep{johnstone2017gaussian}.
